---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data

```{r preprocessing}
library(ggplot2)
library(dplyr)
activity_data <- read.csv("/home/soham/activity.csv")
summary(activity_data)
str(activity_data)
dim(activity_data)
activity_data$date <- as.Date(activity_data$date)

```
Library ggplot2 is used for plotting graphs and library dplyr is used for preprocessing. Before preprocessing date column was of character type,it is converted into date data type.


## What is mean total number of steps taken per day?

```{r mean_total_steps}
Total_steps <- activity_data %>%group_by(date)%>%filter(!is.na(steps))%>%summarise(total_steps = sum(steps, na.rm=TRUE))
Total_steps
ggplot(Total_steps,aes(x=total_steps))+geom_histogram()+labs(x="Total Steps",y = "Frequency",title = "Daily Steps")
mean_toatal_steps <-mean(Total_steps$total_steps)
mean_toatal_steps
median_total_steps <- median(Total_steps$total_steps)
median_total_steps

```

To calculate mean total number of steps per day row's having 'steps' Not Available(NA) are not considered.
## What is the average daily activity pattern?
```{r daily_activity_pattern}
Interval <- activity_data %>%group_by(interval)%>%filter(!is.na(steps))%>%summarise(avg_steps = mean(steps))
Interval
ggplot(Interval,aes(x=interval,y = avg_steps))+geom_line(size = 1)
Interval[which.max(Interval$avg_steps),]

```



## Imputing missing values
To impute missing values following stratigy is used :-
1) Average number of steps are calculated for each interval
2) Not Available(NA) values are imputed by respective average number of steps for that particular interval.
```{r missing_values}
sum(is.na(activity_data$steps))

activity_data_2 <- activity_data
nas <- is.na(activity_data_2$steps)
avg_interval <- tapply(activity_data_2$steps,activity_data_2$interval,mean,na.rm = TRUE)
activity_data_2$steps[nas] <- avg_interval[as.character(activity_data_2$interval[nas])]
names(activity_data_2)
sum(is.na(activity_data_2))
activity_data_2<- activity_data_2[, c("date", "interval", "steps")]
head(activity_data_2)
Total_steps_2 <- activity_data_2 %>%group_by(date)%>%filter(!is.na(steps))%>%summarise(total_steps = sum(steps, na.rm=TRUE))
Total_steps_2

ggplot(Total_steps_2,aes(x = total_steps))+geom_histogram(binwidth  = 1000)+labs(x="Total Steps",y = "Frequency",title = "Daily Steps")

mean_toatal_steps_2 <-mean(Total_steps_2$total_steps)
mean_toatal_steps_2
median_total_steps_2 <- median(Total_steps_2$total_steps)
median_total_steps_2


```



## Are there differences in activity patterns between weekdays and weekends?

```{r activity_pattern}
activity_data_2 <- activity_data_2%>%mutate(week_type = ifelse(weekdays(activity_data_2$date)=="Saturday"|weekdays(activity_data_2$date)=="Sunday","Weekend","Weekday"))

head(activity_data_2)

Interval_2 <- activity_data_2 %>%group_by(interval,week_type)%>%filter(!is.na(steps))%>%summarise(avg_steps_2 = mean(steps))
head(Interval_2)



ggplot(Interval_2, aes(x =interval , y=avg_steps_2, color=week_type)) +
  geom_line() +
  labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps")+facet_wrap(~week_type, ncol = 1, nrow=2) 

```
